\documentclass[../thesis.tex]{subfiles}
\begin{document}

\thispagestyle{empty}

\chapter*{\textbf{Acknowledgements}}
 \addcontentsline{toc}{chapter}{Acknowledgements}

  \begin{onehalfspace} 
Firstly, I would like to express my sincere gratitude to my supervisors, Prof. Daniel Gartner and Prof. Paul Harper for their guidance throughout my research journey. Your support and encouragement has not only helped me grow as a researcher, but also as a person. I could not have imagined having a better supervision team for my PhD.

I would also like to acknowledge the financial support provided by KESS2, along with the Aneurin Bevan University Health Board, which made this research possible. I would like to extend my thanks to the Clinical Futures Team at ABUHB, for contributing their time and knowledge in assisting with this research. I would like to especially thank Dr. Andrew Gray and Jacqui Thornton for your enthusiasm and involvement in this project.

I am grateful to the School of Mathematics staff and my fellow PhD students for creating a supportive community and for making my PhD thoroughly enjoyable. More specifically, I would like to thank Michalis for the endless support, for always having time to fix whatever I had managed to break and most importantly looking after my plant Trev. To Abby and Shauna for always providing a welcome distraction and for our weekly Tuesday lunchtime sessions. To Tim, Michela and Matthew, thank you for your willingness to give up your time to talk, play games and above all, to drink tea.

I would like to thank my friends and family for all your support over the last three and a half years. To my sisters, Sallie and Suzanne, you both inspire me every day, and I will forever be grateful for the love and encouragement you give me. Finally, to my mum; this thesis would not exist without you. Your unconditional love, belief and confidence in me, and words of wisdom is something I will treasure forever.

\end{onehalfspace}
\vfill

% \setcounter{page}{2}
\end{document}
